Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2024-12-10 article.tex
Update literature review and text

%% New math commands
\newcommand{\s}[1]{\overset{*}{#1}}
\newcommand{\RM}{\bm{\Lambda}}
\newcommand{\RMI}{\bm{\Lambda}_0}
\newcommand{\RMT}{\bm{\Lambda}_t}
\newcommand{\RV}{\bm{\psi}}
\newcommand{\magRV}{\psi}
\newcommand{\RMTS}{\s{\bm{\Lambda}}_t}
\newcommand{\bb}{\boldsymbol}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Paper outline:
%
%Intro
%- FV is of interest for CSM
%- Most methods use segregated algorithms, stemming from CFD algorithms
%- Extension to block-coupled Newton methods is not easy, in terms of derivation, and code refactoring (matrix storage, extended stencil, linear solver, etc.).
%- JFNK promises the performance of Newton methods, but without the need to form the full Jacobian. Define JFNK method. Hence, an existing segregated code can easily be adapted to use JFNK without major refactoring, albeit this process is made easier by the availability of open-source JFNK implementations.
%- this paper examines JFNK for linear and nonlinear FV CSM procedures, where the compact stencil approximate Jacobian is used for preconditioning.
Finite volume formulations for solid mechanics are heavily influenced by their fluid mechanics counterparts, favouring segregated implicit and fully explicit methods.
Segregated approaches, where the governing momentum equation is temporarily decomposed into scalar component equations, offer memory efficiency and simplicity of implementation, but the outer coupling Picard iterations often suffer from slow convergence.
Explicit formulations are straightforward to implement and offer superior robustness but are only efficient for high-speed dynamics, where the physics requires small time increments.
In contrast, the finite element community commonly employs Newton-Raphson-type solution algorithms, which necessitate repeated assembly of the Jacobian matrix and solution of the resulting block-coupled non-diagonally dominant linear system.
A disadvantage of traditional Newton-based approaches is that they typically require explicit formation and storage of the Jacobian matrix, which can be computationally expensive and memory-intensive.
A further disadvantage from a finite volume perspective is that extending existing code frameworks from segregated algorithms to a coupled Newton-Raphson-type approach is challenging in terms of the required assembly, storage, and solution of the resulting block-coupled system.
In addition, the derivation of the true Jacobian matrix is non-trivial.
Consequently, similar block-coupled solution finite volume methods are rare in the literature \citep{Das2011, Cardiff2016, Castrillo2024}.
The motivation of the current work is to seek (or exceed) the robustness and efficiency of block-coupled Newton-Raphons approaches in a way that can be easily incorporated into existing segregated solution frameworks.
To this end, the current article examines the efficacy of \emph{Jacobian-free} Newton-Krylov methods, where the quadratic convergence of Newton methods can potentially be achieved without deriving, assembling and storing the exact Jacobian.

Jacobian-free Newton-Krylov methods circumvent the need for the Jacobian matrix by combining the Newton-Raphson method with Krylov subspace iterative linear solvers, such as GMRES, and noticing that such Krylov solvers do not explicitly require the Jacobian matrix.
Instead, only the action of the Jacobian matrix on a solution-type vector is required.
The key step in Jacobian-free Newton-Krylov methods is the approximation of products between the Jacobian matrix and a vector using the finite difference method; that is
\begin{eqnarray}
	\bb{J} \bb{v} \approx \frac{\bb{F}(\bb{x} + \epsilon \bb{v}) - \bb{F}(\bb{x})}{\epsilon}
\end{eqnarray}
where $\mathbf{J}$ is the Jacobian matrix, $\mathbf{x}$ is the current solution vector (e.g. nodal displacements), $\mathbf{v}$ is a vector (e.g., from a Krylov subspace), and $\epsilon$ is a small scalar perturbation.
%Determining the appropriate value for $\epsilon$ requires balancing the truncation error of the finite difference approximation and round-off (numerical precision) error.
With an appropriate choice of $\epsilon$ (balancing truncation and round-off errors), the characteristic quadratic convergence of Newton methods can be achieved without the Jacobian, hence the modifier \emph{Jacobian-free}.
This approach promises significant memory savings over Jacobian-based methods, especially for large-scale, but also potentially for execution time, with appropriate choice of solution components.

A crucial aspect of ensuring the efficiency and robustness of the Jacobian-free Newton-Krylov method is the choice of a suitable preconditioner for the Krylov iterations.
This preconditioner is often derived from the exact Jacobian matrix in traditional Newton methods.
However, the Jacobian-free approach does not allow direct access to the full Jacobian matrix, necessitating an alternative strategy to approximate its action.
To this end, and to extend existing segregated frameworks, we propose using a compact-stencil approximate Jacobian as the preconditioner. This approximate Jacobian corresponds to the matrix typically employed in segregated approaches; similar approaches are successful in fluid mechanics applications \citep{Nishikawa2020, nonNewtonianJFNKPaper}; however, it is unclear if such an approach is suitable for solid mechanics - a question which we hope to answer in this work.
By leveraging this compact-stencil approximate Jacobian, we aim to effectively precondition the Krylov iterations, enhancing convergence while maintaining the memory and computational savings that define the Jacobian-free and segregated methods.
Similarly, if such an approach is efficient, it would naturally fit into existing segregated frameworks, as existing matrix storage and assembly can be reused.

The remainder of the paper is structured as follows:
Section 2 summarises a typical solid mechanics mathematical model and its cell-centred finite volume discretisation.
Section 3 presents the solution algorithms, starting with the classic segregated solution algorithm, followed by the proposed Jacobian-free Newton-Krylov solution algorithm.
The performance of the proposed Jacobian-free Newton-Krylov approach is compared with the segregated approach on several varying benchmark cases in Section 4, where the effect of several factors are examined, including problem dimension, mesh, material model, nonlinear geometry, choice of preconditioner, and other solution parameter.
Finally, the article ends with a summary of the main conclusions of the work.

@@

In this work, we restrict our interest to Lagrangian formulations of the conservation of linear momentum.
Assuming small strains, the linear geometry formulation is expressed in strong integral form as:
\begin{eqnarray} \label{eqn:momentum_lingeom}
    \int_{\Omega} \rho \frac{\partial^2 \bb{u} }{\partial t^2} \, d\Omega
    =
    \oint_{\Gamma} \bb{n} \cdot \bb{\sigma}_s \,  d\Gamma
    + \int_{\Omega}  \bb{f}_b \, d\Omega
\end{eqnarray}
where $\Omega$ is the volume of an arbitrary body bounded by a surface $\Gamma$ with outwards pointing normal $\bb{n}$.
The density is $\rho$, $\bb{u}$ is the displacement vector, $\bb{\sigma}_s$ is the engineering (small strain) stress tensor, and $\bb{f}_b$ is a body force per unit volume, e.g., $\rho \bb{g}$, where $\bb{g}$ is gravity.

More generally,  linear momentum conservation can be expressed in a nonlinear geometry form, which is suitable for finite strains.
Two equivalent nonlinear geometry forms are common: the \emph{total} Lagrangian form:
\begin{eqnarray} \label{eqn:momentum_TL}
    \int_{\Omega_o} \rho_o \frac{\partial^2 \bb{u} }{\partial t^2} d\Omega_o
    =
    \oint_{\Gamma_o} \left( J \bb{F}^{-T} \cdot \bb{n}_o \right) \cdot \bb{\sigma} \ d\Gamma_o
    + \int_{\Omega_o}  \bb{f}_b \, d\Omega_o
\end{eqnarray}
and the \emph{updated} Lagrangian form:
\begin{eqnarray} \label{eqn:momentum_UL}
    \int_{\Omega_u} \frac{\partial }{\partial t} \left( \rho_u \frac{\partial \bb{u} }{\partial t} \right) d\Omega_u
    = \oint_{\Gamma_u}(j\bb{f}^{-T}\cdot{\bb{n}_u)\cdot \bb{\sigma}}\ d\Gamma_u
    + \int_{\Omega_u}  \bb{f}_b \, d\Omega_u
\end{eqnarray}
where subscript $o$ indicates quantities in the initial reference configuration, and subscript $u$ indicates quantities in the updated configuration.
The true (Cauchy) stress tensor is indicated by $\bb{\sigma}$.

The deformation gradient is defined as $\bb{F} = \textbf{I} + (\bb{\nabla} \bb{u})^T$ and its determinant as $J = \text{det}(\bb{F})$.
Similarly, the \emph{relative} deformation gradient is given in terms of the displacement \emph{increment} as $\bb{f}=\textbf{I} + \left[\bb{\nabla}(\Delta \bb{u}) \right]^T$ and its determinant as $j = \text{det}(\bb{f})$.
The displacement increment is the change in displacement between the current time step and the previous time step when the time interval is discretised into a finite number of steps.

@@

The definition of the engineering stress ($\bb{\sigma}_s$) and true stress ($\bb{\sigma}$) in Equations \ref{eqn:momentum_lingeom}, \ref{eqn:momentum_TL} and \ref{eqn:momentum_UL} is given by a chosen mechanical law, e.g. linear elasticity.
Several mechanical laws are considered in this work, as briefly described in Section \ref{sec:test_cases}.

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Newton-Type Solution Methods}
%%--------------------------------------------------------------------------------------------------------------------%%
To facilitate the comparison between classic segregated solution algorithms and the proposed Jacobian-free Newton-Krylov algorithm, the governing linear momentum conservation (Equations \ref{eqn:momentum_lingeom}, \ref{eqn:momentum_TL} and \ref{eqn:momentum_UL}) is expressed in the general form:
\begin{eqnarray} \label{eqn:residual}
	\mathcal{R}(\bb{u}) = \bb{0}
\end{eqnarray}
where $\mathcal{R}$ represents the \emph{residual} (imbalance) of the equation, which is a function of the primary unknown field.
For example, in the linear geometry case, the residual is given as
\begin{eqnarray}
    \bb{R}(\bb{u})
    \;=\;
    \oint_{\Gamma} \bb{n} \cdot \bb{\sigma}_s(\bb{u}) \,  d\Gamma
    + \int_{\Omega}  \rho \bb{g} \, d\Omega
    -  \int_{\Omega} \rho \frac{\partial^2 \bb{u} }{\partial t^2} \, d\Omega
    \;=\; \bb{0}
\end{eqnarray}
where the dependence of the stress tensor on the solution vector is made explicitly clear: $\bb{\sigma}_s(\bb{u})$.

In Newton-type methods, a Taylor expansion about a current point $\bb{u}_k$ can be used to solve Equation \ref{eqn:residual} \cite{Knoll2004}:
\begin{eqnarray}
	\bb{R}(\bb{u}_{k+1}) = \bb{R}(\bb{u}_{k}) \;+\;  \bb{R}'(\bb{u}_{k}) (\bb{u}_{k+1} - \bb{u}_{k}) \;+\; \text{H.O.T.} = \bb{0}
\end{eqnarray}
Neglecting the higher-order terms ($\text{H.O.T.}$) yields the strict Newton method in terms of an iteration over a sequence of linear systems: 
\begin{eqnarray} \label{eq:NewtonRaphson}
	\bb{J}(\bb{u}_k) \delta \bb{u} &=& -\bb{R}(\bb{u}_n), \notag \\
	\bb{u}_{k+1} &=& \bb{u}_k + s \, \delta \bb{u}, \notag \\
	\quad
	k &=& 0,1,...
%    \label{eq:NewtonRaphsonA}
%    \overbrace{\left[ \frac{\partial \mathcal{R}(\bb{u})}{\partial \bb{u}} \right]_n}^{\mathcal{J}} \Delta \bb{u} = -\mathcal{R}(\bb{u})_n \\
%    \label{eq:NewtonRaphsonB}
%    \bb{u}_{n+1} = \bb{u}_{n} + \alpha \Delta \bb{u}
\end{eqnarray}
where $\bb{J} \equiv \bb{R}'$ is the Jacobian matrix.
Starting the Newton procedure requires the specification of $\bb{u}_0$.
%is iteratively solved by linearisation about the current value of the solution, leading to a linear system and iterative update of the solution vector:
%\begin{eqnarray}
%    \label{eq:NewtonRaphsonA}
%    \overbrace{\left[ \frac{\partial \mathcal{R}(\bb{u})}{\partial \bb{u}} \right]_n}^{\mathcal{J}} \Delta \bb{u} = -\mathcal{R}(\bb{u})_n \\
%    \label{eq:NewtonRaphsonB}
%    \bb{u}_{n+1} = \bb{u}_{n} + \alpha \Delta \bb{u}
%\end{eqnarray}
%where subscript $n$ indicates the outer (Newton) iteration index.
The scalar $s > 0$ can be chosen to improve convergence, for example, using a line search or under-relaxation procedure, and is equal to unity in the classic Newton-Raphson approach.
Iterations are performed over this system until the residual $\bb{R}(\bb{u}_n)$ and solution correction $\delta \bb{u}$ are sufficiently small, with appropriate normalisation.

>>>>

%% New math commands
\newcommand{\s}[1]{\overset{*}{#1}}
\newcommand{\RM}{\bm{\Lambda}}
\newcommand{\RMI}{\bm{\Lambda}_0}
\newcommand{\RMT}{\bm{\Lambda}_t}
\newcommand{\RV}{\bm{\psi}}
\newcommand{\magRV}{\psi}
\newcommand{\RMTS}{\s{\bm{\Lambda}}_t}
\newcommand{\bb}{\boldsymbol}
\usepackage{nicefrac}
\usepackage{tikz}
\usepackage{tikz-3dplot}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Paper outline:
%
%Intro
%- FV is of interest for CSM
%- Most methods use segregated algorithms, stemming from CFD algorithms
%- Extension to block-coupled Newton methods is not easy, in terms of derivation, and code refactoring (matrix storage, extended stencil, linear solver, etc.).
%- JFNK promises the performance of Newton methods, but without the need to form the full Jacobian. Define JFNK method. Hence, an existing segregated code can easily be adapted to use JFNK without major refactoring, albeit this process is made easier by the availability of open-source JFNK implementations.
%- this paper examines JFNK for linear and nonlinear FV CSM procedures, where the compact stencil approximate Jacobian is used for preconditioning.
Finite volume formulations for solid mechanics are heavily influenced by their fluid mechanics counterparts, favouring fully explicit \citep{Trangenstein1991, Kluth2010, Lee2013, Haider2017} or segregated implicit \citep{Demirdzic1988, Fryer1991, Demirdzic1995, Jasak2000, Tukovic2013, Cardiff2017, Tukovic2018, Batistic2022} methods.
Segregated approaches, where the governing momentum equation is temporarily decomposed into scalar component equations, offer memory efficiency and simplicity of implementation, but the outer coupling Picard iterations often suffer from slow convergence.
Explicit formulations are straightforward to implement and offer superior robustness but are only efficient for high-speed dynamics, where the physics requires small time increments.
In contrast, the finite element community commonly employs Newton-Raphson-type solution algorithms, which necessitate repeated assembly of the Jacobian matrix and solution of the resulting block-coupled non-diagonally dominant linear system.
A disadvantage of traditional Newton-based approaches is that they typically require explicit formation and storage of the Jacobian matrix, which can be computationally expensive and memory-intensive.
A further disadvantage from a finite volume perspective is that extending existing code frameworks from segregated algorithms to coupled Newton-Raphson-type approaches is challenging in terms of the required assembly, storage, and solution of the resulting block-coupled system.
In addition, the derivation of the true Jacobian matrix is non-trivial.
Consequently, similar block-coupled solution finite volume methods are rare in the literature \citep{Das2011, Cardiff2016, Castrillo2024}.
The motivation of the current work is to seek (or exceed) the robustness and efficiency of block-coupled Newton-Raphons approaches in a way that can be easily incorporated into existing segregated solution frameworks.
To this end, the current article examines the efficacy of \emph{Jacobian-free} Newton-Krylov methods, where the quadratic convergence of Newton methods can potentially be achieved without deriving, assembling and storing the exact Jacobian.

Jacobian-free Newton-Krylov methods circumvent the need for the Jacobian matrix by combining the Newton-Raphson method with Krylov subspace iterative linear solvers, such as GMRES, and noticing that such Krylov solvers do not explicitly require the Jacobian matrix.
Instead, only the action of the Jacobian matrix on a solution-type vector is required.
The key step in Jacobian-free Newton-Krylov methods is the approximation of products between the Jacobian matrix and a vector using the finite difference method; that is
\begin{eqnarray}
	\bb{J} \bb{v} \approx \frac{\bb{F}(\bb{x} + \epsilon \bb{v}) - \bb{F}(\bb{x})}{\epsilon}
\end{eqnarray}
where $\mathbf{J}$ is the Jacobian matrix, $\mathbf{x}$ is the current solution vector (e.g. nodal displacements), $\mathbf{v}$ is a vector (e.g., from a Krylov subspace), and $\epsilon$ is a small scalar perturbation.
%Determining the appropriate value for $\epsilon$ requires balancing the truncation error of the finite difference approximation and round-off (numerical precision) error.
With an appropriate choice of $\epsilon$ (balancing truncation and round-off errors), the characteristic quadratic convergence of Newton methods can be achieved without the Jacobian, hence the modifier \emph{Jacobian-free}.
This approach promises significant memory savings over Jacobian-based methods, especially for large-scale problems, but also potentially for execution time, with appropriate choice of solution components.

A crucial aspect of ensuring the efficiency and robustness of the Jacobian-free Newton-Krylov method is the choice of a suitable preconditioner for the Krylov iterations.
This preconditioner is often derived from the exact Jacobian matrix in traditional Newton methods.
However, the Jacobian-free approach does not allow direct access to the full Jacobian matrix, necessitating an alternative strategy to approximate its action.
To this end, and to extend existing segregated frameworks, this work proposes using a compact-stencil \emph{approximate} Jacobian as the preconditioner. This approximate Jacobian corresponds to the matrix typically employed in segregated solid mechanics approaches; similar approaches are successful in fluid mechanics applications \citep{Mchugh1994, Qin2000, Geuzaine2001, Pernice2001, Knoll2004, Nejat2008, Vaassen2008, Lucas2010, Nejat2011, Nishikawa2020}; however, it is unclear if such an approach is suitable for solid mechanics - a question this work aims to answer.
By leveraging this compact-stencil approximate Jacobian, it is aimed to effectively precondition the Krylov iterations, enhancing convergence while maintaining the memory and computational savings that define the Jacobian-free and segregated methods.
Similarly, if such an approach is efficient, it would naturally fit into existing segregated frameworks, as existing matrix storage and assembly can be reused.

%History:
%Knoll: "The difference in predominating approach (Newton versus Picard) seems stronger for implicit IVPs. Again, the applied mathematics community has focused on Newton-based methods and on converging the nonlinear residual within a time step. In the computational physics community, operator splitting (time splitting, fractional step methods) [20,136] has been the ‘‘bread and butter’’ approach, with little attention to monitoring or converging the nonlinear residual within a time step, often allowing a splitting error to remain in time"
% can be traced back to publications motivated by the solution of ODEs \citep{Brown1986} \citep{Gear1983} and publications motivated by the solution of PDEs \citep{Brown1990} ,\citep{Chen1984}. The primary motivation in all cases appears to be the ability to perform a Newton iteration without forming the Ja- cobian. Within the ODE community these methods helped to promote the use of higher-order implicit integration. The studies on PDE problems focused on the use of nonlinear preconditioning, preconditioners constructed from linear parts of the PDEs, and the addition of globalization methods.

As noted by \citet{Knoll2004}, Jacobian-Free Newton–Krylov methods first appeared in the 1980s and early 1990s for the solution of ordinary and partial differential equations \citep{Gear1983, Chan1984, Brown1986, Brown1990}.
\citet{Mchugh1994} and co-workers demonstrated the potential of Jacobian-free Newton-Krylov approaches for the solution of steady, incompressible, Navier-Stokes problems using a staggered finite volume formulation.
They found the GMRES linear solver to be faster than a conjugate solver, however, the true Jacobian matrix was still evalauted by finite differencing in the construction of the preconditioning matrix.
%%on a uniform but staggered mesh.
%Compared several Krylov linear solvers including CG and GMRes. GMRES found to be best.
%Implementation is simplified using a numerically evaluated Jacobian, while robustness is improved using mesh sequencing and a damped Newton iteration.
%ILU(0) only.
%"they still evaluate the Jacobian matrix for generating the preconditioning matrix and store the matrix in the form of an incomplete lower–upper (ILU) decomposition" from Qin2000.
%Later work, such as the approach by 
\citet{Qin2000} later extended Jacobian-Free Newton–Krylov methods to unsteady compressible Reynolds-averaged Navier–Stokes equations.
% with the Baldwin–Lomax algebraic turbulence model.
Once again it was found that GMRES was more robust than a conjugate gradient linear solver.
For preconditioning, a fully matrix-free approach was proposed, based on the approximate factorization procedure of \citet{Badcock1996}.
%; at each time step is solved iteratively using a Newton/GMRES method; An effective matrix-free preconditioner is proposed to fully avoid matrix storage. GMRES(5 or 10).
%They found GMRES to be more robust than CG. They proposed a matrix-free preconditioner so that the overall method was truly matrix-free, not just Jacobian-free.
\citet{Geuzaine2001} examined the use of lower and higher-order discretisations for the preconditioning matrix for steady, compressible high Reynolds number flows.
% around two-dimensional geometries, with PDE turbulence models: the one-equation Spalart– Allmaras model and the two-equation k–omega model.
They found that the lower-order compact stencil preconditioner gave the best performance when combined with the GMRES linear solver and ILU(k) preconditioner.
As is common with other Newton approaches, a pseudo transient algorithm and mesh sequencing strategy were used to improve the convergence.
%CC FVM.
%pseudotransient approach to improve convergence and mesh sequencing.
%time step is increased inversely with the Newton residual reduction.
%GMRES.
%The baseline preconditioning technique used in this study is an ILU(0) factorization of the distance-1 Jacobian matrix. This matrix is based on a lower-order discretization of the equations because it involves only the edge and vertex neighbors. For comparison purposes, more complex ILU(k ) factorizations and the distance-2 Jacobian matrix, namely, the matrix based on the higher-order discretization.
%Use mesh sequencing strategy.
%GMRES(50). ILU(2) was optimal is one case. They compare a compact and larger preconditioning matrix stencil and the compact (distance-1) method was faster.
With the aim of exploiting mature segregated solution approaches, \citet{Pernice2001} recast the SIMPLE segregated algorithm as a method which operators on residuals, allowing it to act as the preconditioner for a Jacobian-free Newton-Krylov approach.
It was found that the SIMPLE-preconditioned Jacobian-free Newton-Krylov substantially accelerated convergence for the incompressible Navier–Stokes problems examined.
%as preconditioners, they use classical pressure-correction methods such as SIMPLE/SIMPLEC and SIMPLER fulfill this requirement.
GMRES was the adopted linear solver, where the importance of choosing an appropriate restart parameter was noted.
They also found a multigrid approach to be more efficient than ILU for preconditioning, where the implementation was based on the PETSc \citep{PETSc} package.
% as a preconditioner method. MG is more memory efficient than ILU.
%Staggered grid.
%They use Algorithm Inexact Newton Backtracking for globalisation => seems to be a form of line search.
%SIMPLEC used as solve momentum with 5 sweeps of Gauss-Seidel and pressure with 20 sweeps. SIMPLEC is recast as a method which operators on residuals so it can be used sa a preconditioner.
%NK-MG methods with pressure-correction smoothers substantially accelerate the convergence of the pressure-correction methods, albeit at an increase in storage cost, and their performance scales well as problem size is increased.

Over the subsequent two decades, Jacobian-free Newton-Krylov methods have seen increasing application in the field of finite volume computational fluid dynamics, albeit they are still far from widespread.
The use of low-order approximate Jacobians for preconditioning has been established as an efficient and robust approach, e.g. \citep{Nejat2008, Vaassen2008, Nejat2011, Nishikawa2020}.
A particularly attractive application of the low-order-discretisation preconditioned Newton-Krylov is for higher-order discretisations; that is, spatial discretisation of order greater than two.
\citet{Nejat2008} demonstrated such a higher order approach for steady, inviscid compressible flows, with up to fourth order accuracy.
GMRES was the adopted linear solver with the ILU(k) preconditioner and fill-in values (k) ranging from 2 to 4. 
%low-order Jacobian matrix captures the essential physical information about the flow, is narrowly banded after reordering and is much better conditioned than the high-order Jacobian. approximate analytic Jacobian of the first-order flux integral; finite difference approximation to the first-order flux Jacobian; pseudo time step in start-up phase and then switch to Newton; gmres no restart ILU precon (in fill 2-4)
%steady-state computation of inviscid compressible flows.
%high order
\citet{Nejat2011} later applied the second-order variant of the approach to non-Newtonian flows and found Newton-GMRES with ILU(1) to be the most efficient combination.
%The performance of the Newton-BiCGSTAB and the Newton-GMRES solvers are compared with each other. The results show, the ILU(1)/Newton-GMRES is the most efficient combination that is robust even in high Reynolds number shear-thinning fluid flow cases.
Like \citet{Vaassen2008} and many other authors, Nejat and co-workers \citep{Nejat2008, Nejat2011} proposed a special start-up phase to encourage the solution to efficiently reach the Newton method's domain of quadratic convergence.
The use of a pseudo-transient approach combined with an adaptive time step (e.g., based on the convergence of the Newton iterations) is common.
%\citet{Vaassen2008} A pseudo-transient at the beginning, where the time step is computed according to the switched evolution relaxation (SER) (increases the time step inversely to the residual norm reduction) - they use an approximate Jacobian for the preconditioner.
A further promotion for the benefit of Jacobian-free Newton-Krylov methods in fluid dynamics was provided by \citet{Lucas2010}, who demonstrated order of magnitude speedups over nonlinear multigrid methods.
A particular benefit was the insensitivity of the Jacobian-free Newton-Krylov methods to changes in mesh aspect ratio, density, time step and Reynolds number.
%enables a speed up of a factor of 10 compared to standard nonlinear multigrid for two-dimensional, large Reynolds number, unsteady flow computations.
%stiffness induced by the maximum aspect ratio, the grid density, the physical time step and the Reynolds number. Compared to standard nonlinear multigrid, speed ups up to a factor of 25 are achieved.
%Compared to nonlinear mul- tigrid, JFNK is virtually insensitive to changes in these stiffness parameters, hence making JFNK a very robust solution method.
%ILU(k) with k = 1-3. Approximate Jacobian.
%\citet{Nishikawa2020} JFNK with generalized conjugate residual linear solver. Approximate Jacobian for preconditioning -> variable preconditioner. Compared with a simple implicit defect-correction approach with first-order Jacobian and Gauss-Seidel linear solver.
%JFNK is faster and uses less CPU time than IDC.
% NOT JFNK: \citet{Nishikawa2019} pseudo time with adaptive time-step based on the NR residuals. Linear systems solved by Gauss-Seidel relaxation scheme.
A final relevant point is the effect of numerical stabilisation (damping) on the convergence: \citet{Nishikawa2017} demonstrated that a Jacobian-free Newton-Krylov approach was less sensitive to the magnitude of the damping and remained stable for lower values of damping compared with a segregated approach.
% for the alpha-damping scheme when used with a .
%It was found that appropriate choice

Although they are increasingly seen in finite volume computational fluid dynamics, Jacobian-free Newton-Krylov methods have yet to be used for finite volume solid mechanics.
This article aims to address this point by assessing the efficacy of Jacobian-free Newton-Krylov cell-centred finite volume formulations applied to static, dynamic, linear and nonlinear solid mechanics problems.
Particular focus is given to the use of a compact-stencil approximate Jacobian preconditioner, inspired from existing segregated solid mechanics procedures, and ease of implementation into an existing segregated finite volume framework -- OpenFOAM \citep{Weller1998}.
%Novel in this paper:
%- JKNK for FVM CSM/CSD
%- focus on easy-of-implementation in an existing segregated code framework.
%- explore compact approximate Jacobian as a suitable preconditioner for linear and nonlinear solid mechanics;
%- explore multigrid as a preconditioner for JFNK
The remainder of the paper is structured as follows:
Section 2 summarises a typical solid mechanics mathematical model and its cell-centred finite volume discretisation.
Section 3 presents the solution algorithms, starting with the classic segregated solution algorithm, followed by the proposed Jacobian-free Newton-Krylov solution algorithm.
The performance of the proposed Jacobian-free Newton-Krylov approach is compared with the segregated approach on several varying benchmark cases in Section 4, where the effect of several factors are examined, including problem dimension, mesh, material model, nonlinear geometry, choice of preconditioner, and other solution parameter.
Finally, the article ends with a summary of the main conclusions of the work.

@@

In this work, interest is restricted to Lagrangian formulations of the conservation of linear momentum.
Assuming small strains, the linear geometry formulation is expressed in strong integral form as:
\begin{eqnarray} \label{eqn:momentum_lingeom}
    \int_{\Omega} \rho \frac{\partial^2 \bb{u} }{\partial t^2} \, d\Omega
    =
    \oint_{\Gamma} \bb{n} \cdot \bb{\sigma}_s \,  d\Gamma
    + \int_{\Omega}  \bb{f}_b \, d\Omega
\end{eqnarray}
where $\Omega$ is the volume of an arbitrary body bounded by a surface $\Gamma$ with outwards pointing normal $\bb{n}$.
The density is $\rho$, $\bb{u}$ is the displacement vector, $\bb{\sigma}_s$ is the engineering (small strain) stress tensor, and $\bb{f}_b$ is a body force per unit volume, e.g., $\rho \bb{g}$, where $\bb{g}$ is gravity.

More generally,  linear momentum conservation can be expressed in a nonlinear geometry form, which is suitable for finite strains.
Two equivalent nonlinear geometry forms are common: the \emph{total} Lagrangian form:
\begin{eqnarray} \label{eqn:momentum_TL}
    \int_{\Omega_o} \rho_o \frac{\partial^2 \bb{u} }{\partial t^2} d\Omega_o
    =
    \oint_{\Gamma_o} \left( J \bb{F}^{-T} \cdot \bb{n}_o \right) \cdot \bb{\sigma} \ d\Gamma_o
    + \int_{\Omega_o}  \bb{f}_b \, d\Omega_o
\end{eqnarray}
and the \emph{updated} Lagrangian form:
\begin{eqnarray} \label{eqn:momentum_UL}
    \int_{\Omega_u} \frac{\partial }{\partial t} \left( \rho_u \frac{\partial \bb{u} }{\partial t} \right) d\Omega_u
    = \oint_{\Gamma_u}(j\bb{f}^{-T}\cdot{\bb{n}_u)\cdot \bb{\sigma}}\ d\Gamma_u
    + \int_{\Omega_u}  \bb{f}_b \, d\Omega_u
\end{eqnarray}
where subscript $o$ indicates quantities in the initial reference configuration, and subscript $u$ indicates quantities in the updated configuration.
The true (Cauchy) stress tensor is indicated by $\bb{\sigma}$.
The deformation gradient is defined as $\bb{F} = \textbf{I} + (\bb{\nabla} \bb{u})^T$ and its determinant as $J = \text{det}(\bb{F})$.
Similarly, the \emph{relative} deformation gradient is given in terms of the displacement \emph{increment} as $\bb{f}=\textbf{I} + \left[\bb{\nabla}(\Delta \bb{u}) \right]^T$ and its determinant as $j = \text{det}(\bb{f})$.
The displacement increment is the change in displacement between the current time step and the previous time step when the time interval is discretised into a finite number of steps.

@@

The definition of the engineering stress ($\bb{\sigma}_s$) and true stress ($\bb{\sigma}$) in Equations \ref{eqn:momentum_lingeom}, \ref{eqn:momentum_TL} and \ref{eqn:momentum_UL} is given by a chosen mechanical law, e.g. linear elasticity, neo-Hookean hyperelasticity.
Several mechanical laws are considered in this work, as briefly described in Section \ref{sec:test_cases}.

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Newton-Type Solution Methods}
%%--------------------------------------------------------------------------------------------------------------------%%
To facilitate the comparison between the classic segregated solution algorithm and the proposed Jacobian-free Newton-Krylov algorithm, the governing linear momentum conservation (Equations \ref{eqn:momentum_lingeom}, \ref{eqn:momentum_TL} and \ref{eqn:momentum_UL}) is expressed in the general form:
\begin{eqnarray} \label{eqn:residual}
	    \bb{R}(\bb{u}) = \bb{0}
\end{eqnarray}
where $\bb{R}$ represents the \emph{residual} (imbalance) of the equation, which is a function of the primary unknown field.
For example, in the linear geometry case, the residual is given as
\begin{eqnarray}
    \bb{R}(\bb{u})
    \;=\;
    \oint_{\Gamma} \bb{n} \cdot \bb{\sigma}_s(\bb{u}) \,  d\Gamma
    + \int_{\Omega}  \rho \bb{g} \, d\Omega
    -  \int_{\Omega} \rho \frac{\partial^2 \bb{u} }{\partial t^2} \, d\Omega
    \;=\; \bb{0}
\end{eqnarray}
where the dependence of the stress tensor on the solution vector is made explicitly clear: $\bb{\sigma}_s(\bb{u})$.

In Newton-type methods, a Taylor expansion about a current point $\bb{u}_k$ can be used to solve Equation \ref{eqn:residual} \cite{Knoll2004}:
\begin{eqnarray}
	\bb{R}(\bb{u}_{k+1}) = \bb{R}(\bb{u}_{k}) \;+\;  \bb{R}'(\bb{u}_{k}) (\bb{u}_{k+1} - \bb{u}_{k}) \;+\; \text{H.O.T.} = \bb{0}
\end{eqnarray}
Neglecting the higher-order terms ($\text{H.O.T.}$) yields the strict Newton method in terms of an iteration over a sequence of linear systems: 
\begin{eqnarray} \label{eq:NewtonRaphson}
	\bb{J}(\bb{u}_k) \delta \bb{u} &=& -\bb{R}(\bb{u}_n), \notag \\
	\bb{u}_{k+1} &=& \bb{u}_k + s \, \delta \bb{u}, \notag \\
	\quad
	k &=& 0,1,...
%    \label{eq:NewtonRaphsonA}
%    \overbrace{\left[ \frac{\partial \mathcal{R}(\bb{u})}{\partial \bb{u}} \right]_n}^{\mathcal{J}} \Delta \bb{u} = -\mathcal{R}(\bb{u})_n \\
%    \label{eq:NewtonRaphsonB}
%    \bb{u}_{n+1} = \bb{u}_{n} + \alpha \Delta \bb{u}
\end{eqnarray}
where $\bb{J} \equiv \bb{R}' \equiv \partial \bb{R}/\partial \bb{u}$ is the Jacobian matrix.
Starting the Newton procedure requires the specification of $\bb{u}_0$.
%is iteratively solved by linearisation about the current value of the solution, leading to a linear system and iterative update of the solution vector:
%\begin{eqnarray}
%    \label{eq:NewtonRaphsonA}
%    \overbrace{\left[ \frac{\partial \mathcal{R}(\bb{u})}{\partial \bb{u}} \right]_n}^{\mathcal{J}} \Delta \bb{u} = -\mathcal{R}(\bb{u})_n \\
%    \label{eq:NewtonRaphsonB}
%    \bb{u}_{n+1} = \bb{u}_{n} + \alpha \Delta \bb{u}
%\end{eqnarray}
%where subscript $n$ indicates the outer (Newton) iteration index.
The scalar $s > 0$ can be chosen to improve convergence, for example, using a line search or under-relaxation/damping procedure, and is equal to unity in the classic Newton-Raphson approach.
Iterations are performed over this system until the residual $\bb{R}(\bb{u}_n)$ and solution correction $\delta \bb{u}$ are sufficiently small, with appropriate normalisation.
